
```{r}
datadir <- "~/lab/chunks/heatshockStephensProject/data/"
mdatar <- read.csv(paste0(datadir,"countedDataForAnalysissne26.csv"))
datar <- dcast(subset(mdatar,Tag=="UP"),Strain~SampleNum+whichRun+Q+time+replicate,value.var="Counts")
rownames(datar) <- datar[,1]
datar <- datar[,-1]
```

That read in the data. If you are not Darach and you're running this,
you'll have to get your data in somewhere else.


How do the counts look? So we're looking across each row of genes,
and we're asking how many counts do we have?

```{r}
plot(density(apply(datar,1,function(x){sum(x)}),0.1))
```

Basically, I looped between this and looking at the mean vs var
plot from limma, and picked this cut off

```{r}
plot(density(apply(datar,1,function(x){sum(x)})[apply(datar,1,function(x){sum(x)})>15],0.1))

sdatar <- datar[apply(datar,1,function(x){sum(x)})>15,]
```

Here we build the design matrix.

```{r}
d <- dcast(subset(mdatar,Tag=="UP"),
  SampleNum+whichRun+Q+time+replicate~.,value.var="Counts")
shifted <- relevel(factor(!grepl("preShift",d$time)),"FALSE")
shocked <- relevel(factor(grepl("postShock",d$time)),"FALSE")
whichRun <- factor(d$whichRun)
Q <- relevel(factor(d$Q),"FALSE")

library(limma)
design <- model.matrix(~0+whichRun+shifted+shifted:Q+shocked+shocked:Q)
vdatar <- voom(sdatar,design=design,plot=F)
fit <- lmFit(vdatar,design)
```

Then pick contrasts to look at, basically that's a vector of values
that show which weights we're testing for difference associated.

```{r}
data.frame(colnames(design))
```

Which are increased in the second run?
```{r}
runDiff <- eBayes(contrasts.fit(fit,c(-1,1,0,0,0,0,0)))
runDiffTable <- topTable(runDiff,sort.by="logFC",number=10000)
runDiffTable[order(runDiffTable$logFC,decreasing=T),][1:20,]
```

Which are increased in the glutamine flasks regardless of treatment?
```{r}
flaskDiff <- eBayes(contrasts.fit(fit,c(0,0,0,0,1,0,0)))
flaskDiffTable <- topTable(flaskDiff,sort.by="logFC",number=10000)
flaskDiffTable[order(flaskDiffTable$logFC,decreasing=T),][1:20,]
```

Which are increased in shifting? 
```{r}
shiftDiff <- eBayes(contrasts.fit(fit,c(0,0,1,0,0,0,0)))
shiftDiffTable <- topTable(shiftDiff,sort.by="logFC",number=10000)
shiftDiffTable[order(shiftDiffTable$logFC,decreasing=T),][1:20,]
```

Which are increased in glutamine shifting? 
```{r}
qshiftDiff <- eBayes(contrasts.fit(fit,c(0,0,0,0,0,1,0)))
qshiftDiffTable <- topTable(qshiftDiff,sort.by="logFC",number=10000)
qshiftDiffTable[order(qshiftDiffTable$logFC,decreasing=T),][1:20,]
```

Which are increased in shocking? 
```{r}
shockDiff <- eBayes(contrasts.fit(fit,c(0,0,0,1,0,0,0)))
shockDiffTable <- topTable(shockDiff,sort.by="logFC",number=10000)
subset(shockDiffTable[order(shockDiffTable$logFC,decreasing=T),],adj.P.Val<0.05)
```

Which are increased in shocked glutamine flasks? 
```{r}
qshockDiff <- eBayes(contrasts.fit(fit,c(0,0,0,0,0,0,1)))
qshockDiffTable <- topTable(qshockDiff,sort.by="logFC",number=10000)
subset(qshockDiffTable[order(qshockDiffTable$logFC,decreasing=T),],adj.P.Val<0.05)
```


